A Hierarchical Bayesian Framework for Exoplanetary Stellar Obliquity Distribution

### 2023-02-08 src/tex/ms.tex
[showyourwork] overleaf sync

We do not yet understand the importance of stellar inclination in population stellar obliquity inference.
In this work, we propose an approach to infer the stellar obliquity $\psi$ distribution from the sky-projected stellar obliquity $\lambda$ distribution. We show that the stellar obliquity distribution inference weakly depends on stellar inclinations.
In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework to infer the population stellar obliquity distribution.
In Section~\ref{sec:applications}, we apply the framework to simulated data and real observations.
In Section~\ref{sec:jacobian}, we find the mathematical relation between the $\psi$, $\lambda$, and $i_\star$ distributions and explain why stellar inclination is not important in stellar obliquity distribution inference.

@@

To find the unknown stellar obliquity distribution of exoplanetary systems, we build a hierarchical Bayesian framework that takes the sky-projected stellar obliquity $\lambda$ and orbital inclination $i_{\rm orb}$ as observed data. Stellar inclination $i_\star$ is an optional and flexible input. It could be directly taken from previous observations, or if the rotational modulation technique is used, the stellar inclination will be inferred from the stellar rotation period $P_{\rm rot}$, stellar radius $R_\star$, and sky-projected rotational broadening velocity $v\sin{i}_\star$. If the stellar inclination information is not provided, we simply assume an isotropic stellar inclination. As we will show in later sections, assuming an isotropic stellar inclination will not compromise the stellar obliquity distribution inference.

@@

Figure~\ref{fig:graph} shows the probabilistic graphical model for the hierarchical Bayesian framework. We aim to constrain a set of hyperparameters $\bm{\beta}$ that describe the stellar obliquity $\psi$ distribution. The parameter set $\bm{\beta}$ is constrained by $N$ individual systems in which $\psi_n$ is a random variable. $\psi_n$ is computed from the sky-projected stellar obliquity $\lambda_n$, the orbital inclination $i_{{\rm orb},n}$, and the stellar inclination $i_{\star, n}$. 
The parameter $\gamma_{\star, n}$ is a parameter set that contains all the stellar properties other than $i_\star$, e.g., stellar rotation period $P_{\rm rot}$ and radius $R_\star$, if the information is available.
Obs$_{\star,n}$ includes observed data that helps to constrain $i_{\star,n}$ and $\gamma_{\star, n}$.
If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}, where we have $v\sin{i_\star} = 2 \pi R_\star / P_{\rm rot}$.
If the gravity-darkening method is used, the $i_{\star,n}$ is constrained by the anomaly in transit light curves. If asteroseismology is used, the $i_{\star,n}$ is constrained from the periodic variation in the photometric time series.
The sky-projected stellar obliquity $\lambda_n$ can be constrained by observed $\hat{\lambda}_n$ via the Rossiter-McLaughlin effect or gravity darkening.
Lastly, the orbital inclination $i_{{\rm orb},n}$ can be constrained from the transit light curves.

For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to understand if the stellar obliquity is isotropically distributed. If stellar obliquities of misaligned systems follow an isotropic distribution, the $\cos{\psi}$ distribution will be uniformly distributed between -1 to 1. We adopt a two-component mixture Beta distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2. One component is designed for a large population of well-aligned systems, and the other is designed for misaligned systems. The Beta distribution is a flexible distribution that can describe the spike of aligned systems with stellar obliquities close to $0\degr$ or the clustered or broadly distributed misaligned systems.
The probability density function of $\cos{\psi}$ follows 
\begin{align}
    \cos{\psi} &\sim 2\times\Bigl( w_0 {\rm Beta}(a_0, b_0) + w_1 {\rm Beta}(a_1, b_1)\Bigr)-1.
\end{align}
Since the Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\times {\rm Beta} - 1$). For the ${\rm Beta}(a_0,b_0)$ component, we have it to describe the population of aligned systems. For the $a_0/b_0 \gg 1$, the distribution will have a spike at $\cos{\psi} = 1$. For the ${\rm Beta}(a_1,b_1)$ component, we have it to describe the misaligned systems. We design the order of the two components to avoid label switching in the mixture model.
With the design, the hyperparameter and parameter priors are the following:
\begin{align}
    w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
    a_0 &\sim \mathcal{U}(0, 50) \nonumber\\
    b_0 &\sim \mathcal{U}(0, 1) \nonumber\\
    a_1 &\sim \mathcal{U}(0, 10) \nonumber\\
    b_1 &\sim \mathcal{U}(0, 10) \nonumber\\
    \cos{i}_{\star,n} &\sim \mathcal{U}(0, 1) \nonumber\\
    \cos{i}_{{\rm orb},n} &\sim \mathcal{U}(-1, 1) \nonumber\\
    \lambda_{\star,n} &\sim \mathcal{U}(0, \pi).
\end{align}
The likelihood functions follow:
\begin{align}
    \mathcal{L}(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
    \mathcal{L}(i_\star) &\sim \prod_{i=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}).
\end{align}
If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from their observed data.
Here we assume uniform hyperpriors for the Beta distribution. For population inference on a small number of systems (i.e., $N<50$), the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}. When applying the framework to a small sample size, it is important to test the robustness of inferred distributions on different hyperpriors.

The probabilistic model is built with the $\mathtt{PyMC}$ \citep[$\mathtt{v4.1.7}$;][]{pymc} package and the posteriors are sampled using the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, a gradient-based sampling algorithm of the Markov chain Monte Carlo (MCMC). All figures and simulations in this paper are fully reproducible and built with the $\mathtt{showyourwork}$ package. The open source code can be found on \href{https://github.com/jiayindong/polar}{GitHub\,\faGithub}.

@@

\subsection{Simulated Data}

@@

In Figure~\ref{fig:simulation}, we present the inferred stellar obliquity distributions. Since the simulated stellar obliquity distributions only have a single component, we model the data with a single Beta distribution. Each row corresponds to an injected stellar obliquity distribution. The orange curve and contours are the median and uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information, and the blue curve and contours are the ones without stellar inclination information. We simply assume an isotropic distribution of $i_\star$ for these models.
Surprisingly, in all four distributions, the inferred $\cos{\psi}$ distributions with or without $i_\star$ correctly recover the injected distributions shown as grey dashed lines as shown in Figure~\ref{fig:simulation}. Since some of the injected distributions are Normal distributions, it is expected that the inferred distributions, which as Beta distributions, could not exactly describe the injected distributions.

We also examine the importance of orbital inclination to the stellar obliquity inference. Since all the systems we study are transiting exoplanetary systems, we consider an isotropic orbital inclination distribution from $80\degr$ to $90\degr$. This broad inclination range corresponds to an impact parameter range from 0 to 1 for a circular orbit planet with an $a/R_\star$ of 6. We find at this level of orbital inclination deviations, the difference on the inferred stellar obliquity distributions is minimal. For transiting planet systems, approximating orbital inclinations to $90\degr$ will not compromise the stellar obliquity distribution inference.

From the simulated data, we confirm that while stellar inclinations help to constrain the stellar obliquity of individual systems, they are not necessarily required for the population stellar obliquity inference. The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities. The sky-projected stellar obliquity could correctly infer the peak and width of the underlying stellar obliquity distribution.

\begin{figure}[ht!]
    \script{simulation.py}
    \includegraphics[width=\linewidth]{figures/simulation.pdf}
    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with or without stellar inclination information. Each row presents a set of simulated data where the underlying $\cos{\psi}$ distribution is known and shown as grey dashed curves. The inferred $\cos{\psi}$ distributions with or without stellar inclination are shown in orange and blue curves, respectively. The shallow contours are $1\sigma$ and $2\sigma$ uncertainties of the inferred distributions.}
    \label{fig:simulation}
\end{figure}

@@

We then apply the hierarchical Bayesian framework to all exoplanetary systems with sky-projected stellar obliquity $\lambda$ measurements. We use the sample of 161 systems summarized in \cite{Albrecht22} Table A1, which are predominantly Hot Jupiter systems. The inferred $\cos{\psi}$ distribution is shown in Figure~\ref{fig:psi_dist}. The $\cos{\psi}$ distribution peaks at 1 but is nearly flat between $-0.75$ and $0.75$ with no significant clustering. The distribution corresponds to a pile-up of planets with stellar obliquity less than $40\degr$ and an isotropic distribution of planets with stellar obliquity from $40\degr$ to $140\degr$. 
The aligned-system population dominates the distribution with a fraction of $w_0 = 0.67 \pm 0.09$. The hyperposteriors of the Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$.
The misaligned-system population has a fraction of $w_1 = 0.33 \pm 0.09$ and hyperposteriors of $a_1 = 1.56\pm0.93$ and $b_1 = 1.64\pm1.58$.

@@

Our model identifies a pile-up of systems with stellar obliquity less than $40\degr$ and finds that misaligned systems have nearly isotropic stellar obliquity distribution. When comparing the inferred distribution with the prediction of stellar obliquity distributions in different Hot Jupiter origin scenarios, it is in best agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep[e.g.,][]{Nagasawa11, Beague12}. This particular intriguing result should be further examined on more carefully selected Hot Jupiter samples and put constraints on Hot Jupiter origin channels.

@@

In this section, we aim to use mathematical expressions to understand why the stellar obliquity distribution is predominantly determined by the sky-projected stellar obliquity distribution and why the stellar inclination distribution is less important to the inference.

@@

In Figure~\ref{fig:coord}, we build two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ for a given orbital axis $\bf{n}_{\rm orb}$. In both coordinates, we set the observer's line of sight to be one of the two horizontal axes, and the orbital axis of the planet to be the vertical axis. We approximate the orbital inclination of the transiting planet to $90\degr$ here to simplify the problem for a population study. We found in the previous section that the assumption would not compromise the stellar obliquity distribution inference for transiting planets.

@@

In this work, we demonstrate the stellar obliquity distribution inference is predominantly based on the sky-projected stellar obliquity. While stellar inclination improves the stellar obliquity constraint on individual systems, it does not improve the population-level stellar obliquity constraint. 
We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of a population. Information on stellar inclination is not required for the stellar obliquity distribution inference; if the stellar inclination is not provided, we simply assume an isotropic stellar inclination distribution. When combining samples with and without stellar inclination measurements, it is important to check if the $i_\star$ sample is representative of the full sample. If not, since the $i_\star$ sample will constrain stellar obliquities more tightly, the overall distribution could be weighted towards these systems, bias our interpretation.
The hierarchical Bayesian model we introduce in this paper is an open-source code available to the public to apply to a certain sample of targets and customize for different stellar obliquity distributions and priors.

When applying our framework to all exoplanetary systems with sky-projected stellar obliquity measurements, we found no significant clustering near $\cos{\psi} = 0$, at odds with the previous study of systems that have both sky-projected stellar obliquity and stellar obliquity measurements. Potential biases that lead to this discrepancy are open to discussion. 
Their stellar obliquity is nearly isotropically distributed for misaligned exoplanetary systems in support of a multi-planet, convergent-disk-migration history followed by a scattering phase in these planets' dynamical history. Future studies on a more carefully selected Hot Jupiter sample will put better constraints on the origin of Hot Jupiters from their stellar obliquity distribution.

% \section{More}

% First, we apply the framework to the sample of \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ constraints summarized in the \cite{Albrecht22} review. If the host star's rotation period is available, we use the value along with the stellar radii and $v\sin{i_\star}$ to infer the stellar inclination. Otherwise, we use the reported stellar inclination and uncertainty listed in Table A1 of \cite{Albrecht22}.
% The inferred $\cos{\psi}$ distribution is shown in the left panel of Figure~\ref{fig:psi_dist}. The distribution presents a pike at 1, suggesting the majority of exoplanetary systems are well-aligned systems. For misaligned systems, $\cos{\psi}$ presents a cluster near $\cos{\psi} = -0.2$, corresponding to $\psi$ of $100\degr$. The result is consistent with the findings of a preponderance of perpendicular planets in \cite{Albrecht21}.

% To test the dependence of $\cos{\psi}$ distribution on $i_\star$ measurements, we apply the model again to the sample of \numistar systems with both constrained $\lambda$ and $i_\star$. However, in this model, we do not use any additional information on $i_\star$ other than applying an isotropic inclination prior to each system. The corresponding probabilistic graphical model is Figure~\ref{fig:graph} but removing all the dashed edges.
% The $\cos{\psi}$ distribution of the new model is shown in the middle panel of Figure~\ref{fig:psi_dist}. Interestingly, the overall $\cos{\psi}$ distributions of the two models are very similar, despite the fact that the inferred $\psi$ uncertainty for each system is now greater.
% The similarity between the two distributions suggests that, at least for the observed-$i_\star$ sample, the $\cos{\psi}$ distribution mainly depends on the underlying sky-project stellar obliquity $\lambda$ distribution.

% Consequently, we apply our model to all \numall exoplanetary systems that have sky-projected stellar obliquity measurements summarized in \cite{Albrecht22}. The result is presented in the right panel of Figure~\ref{fig:psi_dist}. The $\cos{\psi}$ distribution still spikes at 1 but is nearly flat between $-0.75$ and $0.75$ with no significant clustering. The distribution corresponds to an isotropic distribution of $\psi$ from $40\degr$ to $140\degr$.

% There is no doubt that $i_\star$ measurement on an individual system improves the constraint on $\psi$. However, are $i_\star$ measurements required to constrain the population $\psi$ distribution? Particularly, the left and middle panels in Figure~\ref{fig:psi_dist} show that without $i_\star$ measurements, a similar level of constraint on the $\cos{\psi}$ distribution can be achieved. We perform some calculations and tests to evaluate the necessity of $i_\star$ constraints.

% We also design some $\cos{\psi}$ distributions and infer their distributions purely from the $\lambda$ distributions. From each distribution of $\cos{\psi}$ and assuming a uniform $\theta$, we randomly generate 200 samples of $i_\star$ and $\lambda$ and assume them as \emph{true} values. We then add some Gaussian noises to \emph{true} $\lambda$ and $i_\star$ to simulate \emph{observed} values. We use the uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$ which are typical observation uncertainties. Lastly, from the observed values and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without $i_\star$ measurements.
% We test four different distributions of $\cos{\psi}$: a uniform distribution $\mathcal{U}(-1,1)$ and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
% In all four distributions, we find inferences with or without $i_\star$ are consistent with each other and injected distributions.

% We find the underlying $\lambda$ distributions of the observed-$i_\star$ sample and the full sample by building another hierarchical Bayesian model. We again use a two-component mixture Beta distribution to describe well-aligned and misaligned systems. We extend the Beta distribution support from $[0,1]$ to $[0,\pi]$ by multiplying the distribution by $\pi$. We use the observed $\lambda$ and uncertainties for the likelihoods. Figure~\ref{fig:lam_dist} shows the $\lambda$ distributions of two samples. It clearly shows that in the observed-$i_\star$ sample, misaligned systems are clustered at $\sim 110\degr$, whereas in the full sample, misaligned systems are nearly uniformly distributed.

% The stellar obliquity distribution of exoplanetary systems can be inferred purely from their sky-projected stellar obliquity distribution. When we apply our framework to all systems with $\lambda$ measurements, we find misaligned systems have nearly isotropic stellar obliquity distribution. We compare the inferred distribution from the observation to the prediction of mutual inclination of different eccentricity excitation mechanisms of Hot Jupiters. It is in best agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep{Beague12}.

\vspace{5mm}

>>>>

We do not yet understand the importance of stellar inclination in population stellar obliquity inference.
In this work, we infer the stellar obliquity distribution from sky-projected stellar obliquities and show that the stellar obliquity distribution weakly depends on stellar inclinations.
In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework to infer the population stellar obliquity distribution.
In Section~\ref{sec:applications}, we apply the framework to simulated data and real observations.
In Section~\ref{sec:jacobian}, we find the mathematical relation between the $\psi$, $\lambda$, and $i_\star$ distributions and explain why stellar inclination is not important in stellar obliquity distribution inference.

@@

To find the stellar obliquity distribution of exoplanetary systems, we build a hierarchical Bayesian framework that takes the sky-projected stellar obliquity $\lambda$ and orbital inclination $i_{\rm orb}$ as observed data. Measurements on the stellar inclination $i_\star$ are optional. If the $i_\star$ information is not provided, we simply assume that they follow an isotropic distribution. Otherwise, the observed $i_\star$ could be taken from previous measurements, or if the rotational modulation method is applied, the $i_\star$ could be inferred from the stellar rotation period $P_{\rm rot}$, stellar radius $R_\star$, and sky-projected rotational broadening velocity $v\sin{i}_\star$.

@@

Figure~\ref{fig:graph} shows the probabilistic graphical model for the hierarchical Bayesian framework. We aim to constrain a set of hyperparameters $\bm{\beta}$ that describes the stellar obliquity $\psi$ distribution. The parameter set $\bm{\beta}$ is constrained by $N$ individual systems in which $\psi_n$ is a random variable. $\psi_n$ is computed from the sky-projected stellar obliquity $\lambda_n$, the orbital inclination $i_{{\rm orb},n}$, and the stellar inclination $i_{\star, n}$. 
The parameter $\gamma_{\star, n}$ is a parameter set that contains all the stellar properties other than $i_{\star,n}$, such as the stellar rotation period $P_{{\rm rot},n}$, radius $R_{\star,n}$, and projected rotational velocity $v\sin{i_{\star,n}}$, if they are available.

The sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$ via the Rossiter-McLaughlin effect or gravity darkening.
The orbital inclination $i_{{\rm orb},n}$ is constrained by the transit light curves of exoplanets.
Obs$_{\star,n}$ includes observed data that helps to constrain $i_{\star,n}$ and $\gamma_{\star, n}$.
If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}, where we have $v\sin{i_\star} = 2 \pi R_\star / P_{\rm rot}$.
If the gravity-darkening method is used, the $i_{\star,n}$ could be constrained by the anomaly in transit light curves. If asteroseismology is used, the $i_{\star,n}$ could be constrained from the mode splitting in the Fourier transform of photometric time series.

For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to understand if the stellar obliquity is isotropically distributed. If stellar obliquity distribution follows an isotropic distribution, the $\cos{\psi}$ distribution will be uniformly distributed between -1 to 1. We adopt a two-component mixture Beta distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2. One component is designed for a large population of well-aligned systems, and the other is designed for misaligned systems. The Beta distribution is a flexible distribution that can describe the spike of aligned systems with stellar obliquities close to $0\degr$ or the clustered or broadly distributed misaligned systems.
The probability density function of $\cos{\psi}$ follows 
\begin{align}
    \cos{\psi} &\sim 2\times\Bigl( w_0 {\rm Beta}(a_0, b_0) + w_1 {\rm Beta}(a_1, b_1)\Bigr)-1.
\end{align}
Since the Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\times {\rm Beta} - 1$). For the ${\rm Beta}(a_0,b_0)$ component, we have it to describe the population of aligned systems. For the $a_0/b_0 \gg 1$, the distribution will have a spike at $\cos{\psi} = 1$. For the ${\rm Beta}(a_1,b_1)$ component, we have it to describe the misaligned systems. We design the order of the two components to avoid label switching in the mixture model.
With the design, the hyperparameter and parameter priors are the following:
\begin{align}
    w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
    a_0 &\sim \mathcal{U}(0, 50) \nonumber\\
    b_0 &\sim \mathcal{U}(0, 1) \nonumber\\
    a_1 &\sim \mathcal{U}(0, 10) \nonumber\\
    b_1 &\sim \mathcal{U}(0, 10) \nonumber\\
    \cos{i}_{\star,n} &\sim \mathcal{U}(0, 1) \nonumber\\
    \cos{i}_{{\rm orb},n} &\sim \mathcal{U}(-1, 1) \nonumber\\
    \lambda_{\star,n} &\sim \mathcal{U}(0, \pi).
\end{align}
The likelihood functions follow:
\begin{align}
    \mathcal{L}(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
    \mathcal{L}(i_{\rm orb}) &\sim \prod_{i=1}^N\mathcal{N}(\hat{i}_{{\rm orb},n}, \sigma_{\hat{i}_{{\rm orb},n}}) \nonumber\\
    \mathcal{L}(i_\star) &\sim \prod_{i=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}).
\end{align}
If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from their observed data.
Here we assume uniform hyperpriors for the Beta distribution. For population inference on a small number of systems (i.e., $N \lesssim 50$), the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}. When applying the framework to a small sample size, it is important to test the robustness of inferred distributions on different hyperpriors.

The probabilistic model is built with the $\mathtt{PyMC}$ \citep[$\mathtt{v4.1.7}$;][]{pymc} package and the posteriors are sampled using the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, a gradient-based sampling algorithm of the Markov chain Monte Carlo (MCMC). All figures and simulations in this paper are fully reproducible and built with the $\mathtt{showyourwork}$ package. The open source code can be found on \href{https://github.com/jiayindong/obliquity}{GitHub\,\faGithub\,(https://github.com/jiayindong/obliquity)}.

@@

\subsection{Simulated Data}\label{subsec:sims}

@@

In Figure~\ref{fig:simulation}, we present the inferred stellar obliquity distributions. Since the simulated stellar obliquity distributions only have a single component, we model the data with a single Beta distribution. Each row of Figure~\ref{fig:simulation} corresponds to an injected stellar obliquity distribution. The orange curve and contours are the median and 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information, and the blue curve and contours are the ones without stellar inclination information. We simply assume an isotropic distribution of $i_\star$ for these models.
Surprisingly, in all four distributions, the inferred $\cos{\psi}$ distributions with or without $i_\star$ correctly recover the injected distributions shown as grey dashed lines as shown in Figure~\ref{fig:simulation}. The inferred modes and widths of the stellar obliquity distributions with or without $i_\star$ likelihood are consistent, despite that the inferred distributions without $i_\star$ measurements could have wider uncertainties shown in shallow color contours.
Since the injected distributions are Normal distributions in rows 2--4, it is expected that the inferred distributions, which as Beta distributions, could not exactly describe the injected distributions.

We also examine the importance of orbital inclination $i_{\rm orb}$ to the stellar obliquity distribution inference. Since all the systems we study are transiting-planet systems, we consider an isotropic orbital inclination distribution between $80\degr$ and $90\degr$. This broad inclination range corresponds to an impact parameter range from 0 to 1 with a planet-star separation $a/R_\star$ of 6.
We compare the inferred stellar obliquity distribution by approximating $i_{\rm orb}$ to $90\degr$ with the inferred distribution using the true $i_{\rm orb}$, and find the difference between the two distributions is minimal. For transiting-planet systems, approximating orbital inclinations to $90\degr$ will not compromise the stellar obliquity distribution inference.

From the simulated data, we confirm that while stellar inclinations help to constrain the stellar obliquity of individual systems, they are not necessarily required for the population stellar obliquity inference. The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities. Just from the sky-projected stellar obliquity, we could correctly infer the peak and width of the underlying stellar obliquity distribution. Later in Section~\ref{sec:jacobian}, we derive the Jacobian transformations between $\psi$, $\lambda$, and $i_\star$ to explain why there is no significant information loss when not having $i_\star$ measurements.

\begin{figure}[ht!]
    \script{simulation.py}
    \includegraphics[width=\linewidth]{figures/simulation.pdf}
    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with or without stellar inclination measurements. Each row presents a set of simulated data where the underlying $\cos{\psi}$ distribution is known and shown as grey dashed curves. The inferred $\cos{\psi}$ distributions with or without stellar inclination measurements are shown in orange and blue curves, respectively. The shallow contours are 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
    \label{fig:simulation}
\end{figure}

@@

We next apply the hierarchical Bayesian framework to all exoplanetary systems with sky-projected stellar obliquity $\lambda$ measurements. We use the sample of 161 systems summarized in \cite{Albrecht22} Table A1, which are mainly Hot Jupiter systems. The inferred $\cos{\psi}$ distribution is shown in Figure~\ref{fig:psi_dist}. The $\cos{\psi}$ distribution peaks at 1 but is nearly flat between $-0.75$ and $0.75$ with no significant clustering. The distribution corresponds to a pile-up of planets with stellar obliquity less than $40\degr$ and an isotropic distribution of planets with stellar obliquity from $40\degr$ to $140\degr$. 
The aligned-system population dominates the distribution with a fraction of $w_0 = 0.67 \pm 0.09$. The hyperposteriors of the Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$.
The misaligned-system population has a fraction of $w_1 = 0.33 \pm 0.09$ and hyperposteriors of $a_1 = 1.56\pm0.93$ and $b_1 = 1.64\pm1.58$.

@@

The stellar obliquity distribution suggests that $67\pm9$\% of the systems have stellar obliquity less than $40\degr$, and $33\pm9$\% of the systems have nearly isotropically distributed stellar obliquities spread from $40\degr$ to $140\degr$. The distribution could have important implications for the formation and evolution of close-in planetary systems.
The broad distribution of misaligned systems is in good agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep[e.g.,][]{Nagasawa11, Beague12}, one of the proposed mechanisms that form Hot Jupiters. The intriguing result should be further examined on more carefully selected Hot Jupiter samples and put constraints on Hot Jupiter origin channels.

@@

In this section, we aim to understand why the stellar obliquity distribution is predominantly determined by the sky-projected stellar obliquity distribution and why the stellar inclination distribution is less important to the inference using mathematical expressions.

@@

In Figure~\ref{fig:coord}, we build two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ for a given orbital axis $\bf{n}_{\rm orb}$. In both coordinates, we set the observer's line of sight to be one of the two horizontal axes, and the orbital axis of the planet to be the vertical axis. We approximate the orbital inclination of the transiting planet to $90\degr$ here to simplify the problem for a population study. As discussed in Section~\ref{subsec:sims}, the assumption will not compromise the stellar obliquity distribution inference for transiting planets.

@@

In this work, we demonstrated that the stellar obliquity distribution inference is predominantly determined by the sky-projected stellar obliquities. Without stellar inclination measurements, the stellar obliquity distribution could still be correctly inferred.
We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of transiting-planet systems. Stellar inclination measurement is optional, and if it is not provided, we simply assume an isotropic stellar inclination distribution. The hierarchical Bayesian model is an open-source code that can be found on \href{https://github.com/jiayindong/obliquity}{GitHub\,\faGithub\,(https://github.com/jiayindong/obliquity)}. Users may apply the framework to a certain sample of targets and customize the model for different stellar obliquity distributions and priors.
When jointly modeling the stellar obliquity distribution from one sample with stellar inclination measurements and one without, it is important to check if the $i_\star$ sample is representative of the full sample. If not, since the $i_\star$ sample will constrain stellar obliquities more tightly, the overall distribution could be weighted towards these systems, bias the interpretation.
Lastly, we applied the framework to all exoplanetary systems with sky-projected stellar obliquity measurements and found that $67\pm9$\% of the systems have stellar obliquity less than $40\degr$, and $33\pm9$\% of the systems have nearly isotropically distributed obliquities between $40\degr$ and $140\degr$.
The distribution could have important implications for the formation and evolution of close-in planetary systems and be worthwhile for further investigation.

\vspace*{5mm}
